Numerical simulation of the Cordilleran ice sheet through the last glacial cycle

### 2014-06-07 cordillera-cycle.tex
Polish introduction.

During the last glacial cycle, glaciers and ice caps of the North American
Cordillera were more extensive than today. At the last glacial maximum, a
contiguous blanket of ice extended from Denali Peak in the north to the Puget
Sound in the south \aref. This ice mass is known as the former Cordilleran ice
sheet (Fig.~\ref{fig:locmap}).
\note{Make sure these locations are marked on Fig.~\ref{fig:locmap}.}

For more than a century, exploration and geological investigation of the
Cordillera have led to a large collection of evidence of the former ice cover \aref. This
evidence consists of mapped boundaries of the former ice extent \aref, directions of
past sliding velocities {\aref} and locations of former melt-water streams \aref, as well as
the timing of glacier retreat identified from radiocarbon and cosmogenic dating \aref.

Field-based evidence allowed to reconstruct, with a high degree of confidence,
the maximal extent attained by the Cordilleran ice sheet during the last glacial
cycle \aref. However, former ice thickness and the ice sheet's contribution to the
last glacial maximum sea-level low-stand remain poorly constrained. Moreover,
our understanding of the glaciation history is mainly restricted to the
deglaciation phase and a regional scale, while little is known about the ice
sheet evolution prior to the last glacial maximum extent \aref. Although
time-evolving, whole ice-sheet reconstructions of glacial advance and retreat
patterns are available for the Laurentide and Eurasian ice sheets \aref, such is not
the case for the Cordilleran ice sheet, where complex arrangements of ice flow
directions emanating from multiple glaciation centres have been identified \aref.

@@

Because past climate conditions are subject to considerable uncertainty, our
palaeoclimate forcing is a simplistic approximation of time-evolved temperature
and precipitation fields derived from a combination of a present-day
atmospheric reanalysis (NARR, \aref), lapse-rate corrections and temperature offset
time series. The latter are
obtained by scaling six different palaeo-temperature reconstructions from
proxy records around the globe, including two \chem{\delta^{18}O} records from
Greenland ice cores \aref, two \chem{\delta^{18}O} records from Antarctic ice cores \aref,
and two alkenone unsaturation index records from Northwest Pacific oceanic
sediment cores \aref. Model output is compared to geomorphological evidence in terms
of timing and extent of glaciation and patterns of deglaciation.

@@

% fig:locmap
\begin{figure}
  \includegraphics{locmap}
  \caption{Relief map of northern North America showing a reconstruction of the
           areas once covered by the Cordilleran (CIS), Laurentide (LIS),
           Innuitian (IIS) and Greenland (GIS) ice sheets during the last
           18\,\unit{\chem{^{14}C}\,kyr\,BP} (21.4\,cal\,kyr\,BP)
           \citep{dyke-2004}. The rectangular box denotes the location of the
           modelling domain used in this study. Major mountain ranges covered
           by the ice sheet include the Alaska Range (AR), the Wrangell and
           St.-Elias mountains (WSE), the Selwyn and MacKenzie mountains (SMK),
           the Coast Mountains (CM) and the Rocky Mountains (RM). The background
           map consists of ETOPO1 \citep{data:etopo1} and Natural Earth Data
           \citep{data:naturalearth}.
           \todo{Mark palaeo-ice sheets and mountain ranges on the map.}}
  \label{fig:locmap}
\end{figure}

>>>>

During the last glacial cycle, glaciers and ice caps of the North American
Cordillera were more extensive than today. At the last glacial maximum, a
contiguous blanket of ice extended from the Alaska Range in the north to the
North Cascades in the south \aref. This ice mass is known as the former
Cordilleran ice sheet (Fig.~\ref{fig:locmap}).

For more than a century, exploration and geological investigation of the
Cordillera have led to a large collection of evidence of the former ice cover
\aref. This evidence consists of mapped boundaries of the former ice extent
\aref, directions of past sliding velocities {\aref} and locations of former
melt-water streams \aref, as well as the timing of glaciation identified from
radiocarbon dating \aref, cosmogenic dating \aref, and the offshore sedimental
record \aref.

Field-based evidence allowed to reconstruct, with a high degree of confidence,
the maximal extent attained by the Cordilleran ice sheet during the last glacial
cycle \aref. However, former ice thickness and the ice sheet's contribution to the
last glacial maximum sea-level low-stand remain poorly constrained. Moreover,
our understanding of the glaciation history is mainly restricted to the
deglaciation phase and a regional scale, while little is known about the ice
sheet evolution prior to the last glacial maximum extent \aref. Although
time-evolving, whole ice-sheet reconstructions of glacial advance and retreat
patterns are available for the Laurentide and Eurasian ice sheets \aref, such is not
the case for the Cordilleran ice sheet, where complex arrangements of ice flow
directions emanating from multiple glaciation centres have been identified, but
remain poorly understood \aref.

@@

Because past climate conditions are subject to considerable uncertainty, our
palaeoclimate forcing is a simplistic approximation of time-evolved temperature
and precipitation fields derived from a combination of a present-day
atmospheric reanalysis (NARR, \aref), lapse-rate corrections,
and temperature offset time series. The latter are
obtained by scaling six different palaeo-temperature reconstructions from
proxy records around the globe, including two \chem{\delta^{18}O} records from
Greenland ice cores \aref, two \chem{\delta^{18}O} records from Antarctic ice cores \aref,
and two alkenone unsaturation index records from Northwest Pacific oceanic
sediment cores \aref. Model output is compared to geomorphological evidence in terms
of timing and extent of glaciation and patterns of deglaciation.

@@

% fig:locmap
\begin{figure}
  \includegraphics{locmap}
  \caption{Relief map of northern North America showing a reconstruction of the
           areas once covered by the Cordilleran (CIS), Laurentide (LIS),
           Innuitian (IIS) and Greenland (GIS) ice sheets during the last
           18\,\unit{\chem{^{14}C}\,kyr\,BP} (21.4\,cal\,kyr\,BP)
           \citep{dyke-2004}. The rectangular box denotes the location of the
           modelling domain used in this study. Major mountain ranges covered
           by the ice sheet include the Alaska Range (AR), the Wrangell and
           St.-Elias mountains (WSE), the Selwyn and MacKenzie mountains (SMK),
           the Skeena Mountains, the Coast Mountains (CM), the Rocky Mountains
           (RM) and the North Cascades (NC). The background
           map consists of ETOPO1 \citep{data:etopo1} and Natural Earth Data
           \citep{data:naturalearth}.
           \todo{Mark palaeo-ice sheets and mountain ranges on the map.}}
  \label{fig:locmap}
\end{figure}
